Accessible and Completed Datasets yet Majority are Bronze-Graded and not Updated Regularly with Missing Metadata

### 2025-05-15 paper/paper.qmd
Finished data section

For the remainder of the paper, the data section (@sec-data) looks at the data used and how it was retrieved alongside the characteristics of the data, the data's limitations, and our variables of interest for our analysis. In the results section (@sec-results) we looked at the data more in-depth through graphs. Finally, the appendix (@sec-appendix) includes acknowledgements and any additional information related to the paper.

@@

## Measurement
The Information & Technology department at Open Data Toronto collected datasets on their portal using the CKAN Datastore API
[@citeCatalogue; @citeCatalogueSteps]. Open Data Toronto uses a metric called the "Data Quality Score" in order to give each dataset in their catalogue a grade indicating "bronze", "silver", or "gold", which can be seen through on the webpage for each dataset on Open Data Toronto portal website. In order to create the "Data Quality Score", they assembled the Data Quality Working Group, which consist of a diverse group of people from consumers of datasets to people who create datasets [@citeCatalogue; @citeCatalogueSteps]. Open Data Toronto first reviewed various literature such as academic papers and industry white papers to compile 15 dimensions used to measure quality [@citeCatalogue; @citeCatalogueSteps]. The dimensions selected were as follows: Interpretability ("How easy it is to understand the data?"), Usability ("How easy is it to work with the data?"), Metadata ("Is the data well described?"), Freshness ("How close to creation time is the data published?"), Granularity ("How atomic is the data?"), Completeness ("How much data is missing?"), and Accessibility ("Is the data easy to access?") [@citeCatalogue; @citeCatalogueSteps]. From there the Data Quality Working Group were surveyed and asked to rank the importance of each dimension when it comes to assessing data quality where 1 represents the most important and 7 being the least important. The results from the survey would then help the team determine the weighting of each dimension towards the overall data quality score. Some of the dimensions were combined or removed. For example, granularity and interpretability were removed and interpretability was combined with usability. The team then used the ranking weighting method, Sum and Reciprocal, and obtain the following weights for each dimension alongside the type of questions asked for each dimension [@citeCatalogue; @citeCatalogueSteps]:

![Weighting](other/images/weighting_of_dimensions.png)

For more technical details about how each dimension is calculated for a dataset and the final data quality score, it can be through the following link: [https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb](https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb) [@citeDQSNotebook]. However, it is worth noting that each dimension are calculated based on data obtained on the metadata information and the dataset itself. If the data quality score of a dataset has a normalized score of less than 60%, they are given the grade "Bronze", if the normalized score is between 60% and 80%, it is given the grade "Silver", and finally if the normalized score is over 80%, tit is given the grade "Gold". The score for each dataset in the portal is recalculated every week by the team [@citeCatalogue].

@@

Our variables of interest are "accessibility", "completeness", "freshness", "metadata", "usability", and "grade".

@@

>>>>

For the remainder of the paper, the data section (@sec-data) looks at the data used and how it was retrieved alongside the characteristics of the data, the data's limitations, and our variables of interest for our analysis. In the results section (@sec-results) we looked at the data more in-depth through graphs. With the discussion section (@sec-discussion), we will provide an overview of what was done in our results, discussing our results and its real-world implications, and indicate areas of improvements for our analysis and directions for future works. Finally, the appendix (@sec-appendix) includes acknowledgements and any additional information related to the paper.

@@

## Measurement
The Information & Technology department at Open Data Toronto collected datasets on their portal using the CKAN Datastore API
[@citeCatalogue; @citeCatalogueSteps]. Open Data Toronto uses a metric called the "Data Quality Score" in order to give each dataset in their catalogue a grade indicating "bronze", "silver", or "gold", which can be seen through on the webpage for each dataset on Open Data Toronto portal website. In order to create the "Data Quality Score", they assembled the Data Quality Working Group, which consist of a diverse group of people from consumers of datasets to people who create datasets [@citeCatalogue; @citeCatalogueSteps].

Open Data Toronto first reviewed various literature such as academic papers and industry white papers to compile 15 dimensions used to measure quality [@citeCatalogue; @citeCatalogueSteps]. The dimensions selected were as follows: Interpretability ("How easy it is to understand the data?"), Usability ("How easy is it to work with the data?"), Metadata ("Is the data well described?"), Freshness ("How close to creation time is the data published?"), Granularity ("How atomic is the data?"), Completeness ("How much data is missing?"), and Accessibility ("Is the data easy to access?") [@citeCatalogue; @citeCatalogueSteps].

From there the Data Quality Working Group were surveyed and asked to rank the importance of each dimension when it comes to assessing data quality where 1 represents the most important and 7 being the least important. The results from the survey would then help the team determine the weighting of each dimension towards the overall data quality score. Some of the dimensions were combined or removed. For example, granularity and interpretability were removed and interpretability was combined with usability. The team then used the ranking weighting method, Sum and Reciprocal, and obtain the following weights for each dimension alongside the type of questions asked for each dimension [@citeCatalogue; @citeCatalogueSteps]:

![Weighting](../other/images/weighting_of_dimensions.png)

For more technical details about how each dimension is calculated for a dataset and the final data quality score, it can be through the following link:[https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb](https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb) [@citeDQSNotebook]. However, it is worth noting that each dimension are calculated based on data obtained on the metadata information and the dataset itself. If the data quality score of a dataset has a normalized score of less than 60%, they are given the grade "Bronze", if the normalized score is between 60% and 80%, it is given the grade "Silver", and finally if the normalized score is over 80%, tit is given the grade "Gold". The score for each dataset in the portal is recalculated every week by the team [@citeCatalogue].

@@

Our variables of interest that we used in our analysis are the following: "accessibility", "completeness", "freshness", "metadata", "usability", and "grade". "Accessiblity" is a score from 0 to 1 that indicates the degree that the dataset can be access or not through the Open Data Toronto API, keywords or tags, and automated data pipelines with 1 being that it can not be accessed with the various methods noted and 0 if not at all. "Completeness" is a score from 0 to 1 that indicates how much of the data is missing with 1 being that there is no missing data and 0 being that all data fields are empty or missing. The "freshness" variable is a score from 0 to 1 that indicates how up-to-date the data where a shorter time duration between the recent refresh date and time and the previous refresh date and time before that gives a higher score. The "metadata" variable indicates how complete the following metadata fields are for a dataset from a scale of 0 to 1: Description, Limitations, Topics, Contact Email. The more metadata fields compelted, the higher the metadata score. The "usability" variable is a score from 0 to 1 indicating how easy it would be to use the dataset and this is determine by the proportion of meaningful column names or in other words column names with English words in it.

@@

# Discussion {#sec-discussion}
In @sec-results, we

## Point 1

## Point 2

## Areas of improvement

## Next steps
